---
output: github_document
always_allow_html: yes
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

![build](https://github.com/IMMM-SFA/gamut/workflows/build/badge.svg) [![codecov](https://codecov.io/gh/IMMM-SFA/gamut/branch/master/graph/badge.svg?token=uF3EvxwvCO)](https://codecov.io/gh/IMMM-SFA/gamut) 
[![DOI](https://zenodo.org/badge/203447802.svg)](https://zenodo.org/badge/latestdoi/203447802) 

# `gamut`

#### **G**eospatial **A**nalytics for **M**ultisector **U**rban **T**eleconnections

## Description
`gamut` is a tool for exploring *teleconnections* between cities of the United States and human activities that occur in their associated water supply catchments. A *teleconnection* is a causal connection or correlation between human and environmental phenomena that occur a long distance apart.


```{r,echo=FALSE, message=FALSE, warning=FALSE,fig.width=10,fig.height=5}
library(gamut)
library(ggplot2)
library(sf)
library(st)

# define data directory

your_gamut_data_directory <- "C:/Geospatial_Input_Datasets/"

get_cities() %>%
     subset(key_watershed == TRUE) ->
     city_watershed_mapping

city_watershed_mapping$DVSN_ID %>% unique() -> watersheds
city_watershed_mapping$city_uid %>% unique() -> cities

import_shapefile(paste0(your_gamut_data_directory,
                        "/misc/states_21basic/states.shp")) %>% 
  subset(!STATE_NAME %in% c("Alaska", "Hawaii"))->
  states

import_shapefile(paste0(your_gamut_data_directory,
                        "/water/CWM_v2_2/World_Watershed8.shp")) %>%
                   st_as_sf() %>%
                   subset(DVSN_ID %in% watersheds) ->
                   watersheds
                 
import_shapefile(paste0(your_gamut_data_directory, 
                 "/water/CWM_v2_2/City_Centroid.shp")) %>%
  st_as_sf() %>% subset(City_ID %in% cities) -> intakes

#raster::raster("C:/CERF/Data/MSR_50M/MSR_50M/MSR_50M.tif") -> relief
# crop(relief, states) %>% mask(states) -> relief2
# relief_spdf <- as(relief2, "SpatialPixelsDataFrame")
# relief <- as.data.frame(relief_spdf)


ggplot() +
  # geom_raster(data = relief, aes(x = x,
  #                                y = y,
  #                                alpha = MSR_50M)) +
  # scale_alpha(name = "", range = c(0.6, 0), guide = F)  +
  geom_sf(data=states, fill = "lightgrey", size = 0.7, color = "black") +
  geom_sf(data=watersheds, fill = "steelblue", alpha = 0.4) +
  geom_sf(data=intakes, color = "red4", alpha = 0.9, size = 2) +
  scale_fill_manual(values = "steelblue") +
  coord_sf() +
  theme_void() +
  theme(
    legend.position = c(0.1, 0.15),
    legend.text.align = 0,
    legend.background = element_rect(fill = alpha('white', 0.0)),
    legend.text = element_text(size = 19, hjust = 0, color = "#4e4d47"),
    plot.title = element_text(size = 15, hjust = 0.5, color = "#4e4d47"),
    legend.title = element_blank(),
    plot.margin = unit(c(.5,.5,.2,.5), "cm"),
    panel.spacing = unit(c(-.1,0.2,.2,0.2), "cm"),
    panel.border = element_blank(),
    plot.caption = element_text(size = 6,
                                hjust = 0.92,
                                margin = margin(t = 0.2,
                                                b = 0,
                                                unit = "cm"),
                                color = "#939184")
  ) +
  labs(x = NULL,
       y = NULL,
       title = "Watersheds and Cities")

```

## Get Started with `gamut`
`gamut` can be installed directly from this repository using the R `devtools` package. From an R prompt, run the command:
```r
devtools::install_github('IMMM-SFA/gamut')
```

##### Data Files

The table below shows all the files used within the `gamut` software package.
    
```{r, echo=FALSE,message=FALSE}
library(tibble)
library(knitr)

tribble(
  ~`Teleconnect Name`,	~`Sub Folder Location`, ~`Full File Path`,~`Data Location`,
  "watersheds", "water","water/CWM_v2_2/World_Watershed8.shp","https://knb.ecoinformatics.org/view/doi%3A10.5063%2FF1J67DWR",
  "withdrawal","water","water/CWM_v2_2/Snapped_Withdrawal_Points.shp","https://knb.ecoinformatics.org/view/doi%3A10.5063%2FF1J67DWR",
  "citypoint","water","water/CWM_v2_2/City_Centroid.shp","https://knb.ecoinformatics.org/view/doi%3A10.5063%2FF1J67DWR",
  "powerplants","water","water/UCS-EW3-Energy-Water-Database.xlsx","https://www.ucsusa.org/resources/ucs-ew3-energy-water-database",
  "crop","land","land/2016_90m_cdls/cdl_lowres_usa.img","https://www.nass.usda.gov/Research_and_Science/Cropland/Release/",
  "crop_attributes","land","land/2016_90m_cdls/cdl_lowres_usa.img.vat.dbf","https://www.nass.usda.gov/Research_and_Science/Cropland/Release/",
  "irrigation","land","land/Version2_USA_Demeter.csv","NA",
  "nlud","land","land/usa_nlud_LR.tif","https://drive.google.com/file/d/1vmNfwjcaLf0sZTYJ1wsB3liG37sN8gyC/view",
  "hydro","energy","energy/EHA_Public_PlantFY2019_GIS_6/ORNL_EHAHydroPlant_PublicFY2019final.xlsx","https://hydrosource.ornl.gov/node/250",
  "transfers","NA","NA","NA",
  "climate","land","land/kop_climate_classes.tif","http://koeppen-geiger.vu-wien.ac.at/present.htm",
  "HUC4","water","water/USA_HUC4/huc4_to_huc2.shp","http://prd-tnm.s3-website-us-west-2.amazonaws.com/?prefix=StagedProducts/Hydrography/WBD/National/GDB/",
  "population","land","land/pden2010_block/pden2010_60m.tif","https://www.sciencebase.gov/catalog/item/57753ebee4b07dd077c70868",
  "runoff","water","water/Historical_Mean_Runoff/USA_Mean_Runoff.tif","NA",
  "nhd_flow","water","water/Watershed_Flow_Contributions/UWB_Intake_Flows.shp","NA",
  "contributions","water","water/Watershed_Flow_Contributions/Watershed_Contributions.csv","NA"
) -> files

kable(files)
```

##### Data Downloading Instructions

To download all of the `gamut` input datasets, use the `get_gamut_datasets` function as seen below. This will download all of the necessary data to run the `gamut` package.

```r
get_gamut_datasets("Download Location", "Unzip Location")
```


## Usage

Once all the necessary data is organized in the correct format, the package is ready to be used. The main function is `count_watershed_teleconnections`. Within this function, set up your data directory and select cities. The city names need to be in the format of `City | State`, and for multiple cities, place them inside `c()`.

Here is an example of what you would type into your console: 

```r
count_watershed_teleconnections(data_dir = "your/gamut/data_dir", cities = c("Portland | OR", "Charlotte | NC", "Austin | TX"))
```

The package will cycle through each city and their respective watersheds, and produce a table with several columns of information. To learn what each of these variables mean, scroll to the bottom of this section and see the table of variables. The result of this function will look something like this:

```{r,echo=FALSE, message=FALSE, warning=FALSE, results='hide'}
suppressMessages(count_watershed_teleconnections(your_gamut_data_directory,
                                                 cities = c("Portland | OR", "Knoxville | TN"))) -> results
gsub("\\|", ", ", results$city) -> results$city
gsub("\\ ", "", results$city) -> results$city
gsub("\\,", ", ", results$city) -> results$city
```

```{r,echo=FALSE, message=FALSE, warning=FALSE}
knitr::kable(results)
```

This table can be used to compare different variables between multiple cities. Below is a graph comparing how much developed land are in cities' watersheds. 

```{r, echo=FALSE,message=FALSE}
results %>% 
  ggplot(aes(x = reorder(city, developed_fraction), y = developed_fraction)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  coord_flip() +
  labs(x = NULL, y = "Fraction of Developed Land") +
  theme_classic() 
```


The table below shows explanations for each of these variables that are created through this function:

```{r, echo=FALSE,message=FALSE}
tibble::tribble(
  ~`Variable Name`,	~Description, ~Units,
  "city_population", "The population of the city being analyzed","people",
  "n_watersheds","Number of watersheds that city uses to source drinking water","watersheds",
  "n_other_cities","Number of other cities pulling off the same watersheds","cities",
  "dependent_city_pop","Total population of people dependent on that city's watersheds","people",
  "watershed_area_sqkm","Combined area of all the source watersheds of a city","square kilometers",
  "storage_BCM","Combined storage capacity of all the city catchments","billion cubic meters",
  "yield_BCM","Combined yield capacity of all the city catchments","billion cubic meters",
  "irr_cons_BCM","Combined water consumption that is used for irrigation with the watersheds","billion cubic meters",
  "n_climate_zones","Number of climate zones that the source watersheds cover","zones",
  "n_transfers_in","Number of interbasin transfers that flow into the source watersheds","transfers",
  "n_transfers_out","Number of interbasin transfers that flow out of the source watersheds","transfers",
  "n_transfers_within","Number of water transfers that occur within the source watersheds","transfers",
  "n_hydro_plants","Number of hydro electric power plants operating within the source watersheds","plants",
  "n_thermal_plants","Number of thermal power plants operating within the source watersheds","plants",
  "n_fac_agcrop","Number of agricultural crop facilities within the source watersheds","facilities",
  "n_fac_aglivestock","Number of agicultural livestock facilities within the source watersheds","facilities",
  "n_fac_cnsmnf","Number of construction and manufacturing facilities within the source watersheds ","facilities",
  "n_fac_mining","Number of mining facilities within the source watersheds","facilities",
  "n_fac_oilgas","Number of oil and gas facilities within the source watersheds","facilities",
  "n_fac_total","Total number of facilities operating within the source watersheds","facilities",
  "hydro_gen_MWh","Combined hydro electric generation from all the facilities within the source watersheds","Megawatthours",
  "thermal_gen_MWh","Combined thermal generation from all the facilities within the source watersheds","Megawatthours",
  "thermal_cons_BCM","Combined water consumption that is used for thermal generation","billion cubic meters",
  "thermal_with_BCM","Combined water withdrawal for thermal generation","billion cubic meters",
  "n_utilities","Number of electric utilities within the source watersheds","utilities",
  "n_ba","Number of balancing authorities within the source watersheds","balancing authorities",
  "n_crop_classes","Total number of different types of crops within the source watersheds","crops",
  "cropland_fraction","Fraction of land that is used for crops within the source watersheds","NA",
  "developed_fraction","Fraction of land that is developed within the source watersheds","NA",
  "ag_runoff_max","Max amount of agricultural runoff within the source watersheds","",
  "ag_runoff_av_exgw","","",
  "ag_runoff_av","Average runoff from agricultural lands","",
  "dev_runof_max","Max amount of agricultural runoff within the source watersheds","",
  "dev_runof_av_exgw","","",
  "dev_runof_av","Average runoff from developed lands","",
  "np_runoff_max","Max amount of non-point source runoff within the source watersheds","",
  "np_runoff_av_exgw","","",
  "np_runoff_av_exgw_unweighted","","",
  "np_runoff_av","Average non-point source runoff.","",
  "n_economic_sectors","Total number of different economic sectors within the source watersheds","sectors",
  "max_withdr_dis_km","Maximum distance between a city's intake points","kilometers",
  "avg_withdr_dis_km","Average distance between a city's intake points","kilometers",
  "n_treatment_plants","Total number of waste water treatment plants operating within the source watersheds","plants",
  "watershed_pop","Total number of people living within the source watershed boundaries","people",
  "pop_cons_m3sec","Combined water consumption from the source watersheds that is used for people","m3/sec",
  "av_fl_sur_conc_pct","","%",
  "av_fl_sur_conc_pct_unweighted","","%",
  "av_ro_sur_conc_pct","","%",
  "av_fl_all_conc_pct","","%",
  "av_ro_all_conc_pct","","%",
  "av_fl_max_conc_pct","","%",
  "av_ro_max_conc_pct","","%",
  "surface_contribution_pct","","%",
  "importance_of_worst_watershed_pct","","%",
) -> variables
kable(variables)
```

## Support

For any questions about the package, please contact any of the contributors below:

Sean Turner:
sean.turner@pnnl.gov

Kristian Nelson:
kristian.nelson@pnnl.gov

Chris Vernon
chris.vernon@pnnl.gov


## Authors and Acknowledgement 

Authors: Sean Turner, Kristian Nelson, Chris Vernon, Jennie Rice, Casey Burleyson, Ryan McManamay, Kerim Dickson

This research was supported by the US Department of Energy, Office of Science, as 
part of research in the MultiSector Dynamics, Earth and Environmental System Modeling Program.





